### 2023-01-31 article3.qmd
update

```{r, load libraries}
library(akima)
library(broom)
library(tidyverse)
library(cmocean)
library(lubridate)
library(scales)
library(patchwork)
library(viridis)
library(ggridges)
library(ggvis)
library(pracma)
library(gbm)
library(MLmetrics)
library(purrr)
library(arrow)
library(plotly)
library(oce)
library(ggplot2)
library(latex2exp)
library(cowplot)
library(RColorBrewer)
library(vroom)
library(castr)
library(reactablefmtr)
```

@@

#### When drifting

@@

#### When profiling

@@

```{r, uvp_data on the vertical}
#| label: fig-uvp-vertical-profiles
#| fig-cap: Normalized concentrations of particles for all UVP6 vertical profiles. Plots were arranged to respect the deployment of each couple of floats (e.g. the first two rows of plots are related to co-deployed floats 6904241 and 1902578). The colorbar depicts the log$_{10}$ of normalized concentrations. Grey areas indicate that particles were not measured in the size range of interest.

@@

# order floats by couple of deployment
particles_bis$wmo <- factor(particles_bis$wmo, levels = c(6904241,1902578,6904240,4903634))

@@

## plot absolute concentrations
particles_bis |> filter(depth < 2000) |>
  #group_by(size) |>
  ggplot() + facet_grid(wmo~size) + theme_bw() +
  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
  #geom_point(aes(x=juld, y=depth, colour=rel_conc), size=1, shape=15) +
  scale_colour_viridis(option = 'H', limits = c(-3,-1), oob = scales::squish, na.value = 'grey') +
  #scale_colour_viridis(option = 'H') +
  scale_y_reverse() +
  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
  labs(colour = 'Normalized concentrations (#/L)', x = 'Month', y = 'Depth (m)') +
  #geom_path(data = particles_bis, aes(juld, MLD), colour = "black") +
  theme(legend.position = 'bottom') #+
  #geom_hline(yintercept = c(200,500,1000), linetype = 'dotted') #+
  #guides(colour = guide_legend(title.position = "bottom"))

@@

### Remineralization processes

```{r, monthly anomaly between drifting at 500 and 1000 m}

tmp <- uvp_data |> filter(park_depth != '200 m', month > 5) |> group_by(wmo, size, month, park_depth) |> 
  summarize(monthly_mean_conc = mean(conc), monthly_median_conc = median(conc)) |>
  # compute concentration anomaly at 500 and 1000 m
  mutate(anomaly = monthly_mean_conc - lead(monthly_mean_conc)) |> drop_na(anomaly) |>
  select(wmo, size, month, anomaly) |> mutate(month = month.abb[month]) |> 
  mutate(size = as.character(size)) |>
  #mutate(anomaly = round(anomaly, 3))
  ungroup()

tt <- tmp |> filter(wmo == 6904241) |> select(size, month, anom = anomaly)
#tt <- tt |> pivot_wider(names_from = month, values_from = anom) |> mutate(size = as.character(size))

x <- c('Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec')
y <- unique(tt$size)
data <- expand.grid(X=x, Y=y)
data$Z <- tt$anom

ggplot(data, aes(X, Y, fill= Z)) + 
  geom_tile() + scale_fill_gradientn(colors = c('darkblue', 'blue', 'white', 'red', 'darkred'))

@@

>>>>

```{r, load libraries}
library(akima)
library(broom)
library(tidyverse)
library(cmocean)
library(lubridate)
library(scales)
library(patchwork)
library(viridis)
library(ggridges)
library(ggvis)
library(pracma)
library(gbm)
library(MLmetrics)
library(purrr)
library(arrow)
library(plotly)
library(oce)
library(ggplot2)
library(latex2exp)
library(cowplot)
library(RColorBrewer)
library(vroom)
library(castr)
library(reactablefmtr)
library(gt)
```

@@

#### On drifting mode

@@

#### On profiling mode

@@

```{r, uvp_data on the vertical}
#| label: fig-uvp-vertical-profiles
#| fig-cap: Normalized concentrations of particles for all UVP6 vertical profiles. Plots were arranged to respect the deployment of each couple of floats (e.g. the first two rows of plots are related to co-deployed floats 6904240 and 4903634). The colorbar depicts the log$_{10}$ of normalized concentrations. Grey areas indicate that particles were not measured in the size range of interest.

@@

# order floats by couple of deployment
particles_bis$wmo <- factor(particles_bis$wmo, levels = c(6904240,4903634,6904241,1902578))

@@

## plot absolute concentrations
up_part <- particles_bis |> filter(depth < 2000, size < 512) |>
  #group_by(size) |>
  ggplot() + facet_grid(wmo~size) + theme_bw() +
  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
  #geom_point(aes(x=juld, y=depth, colour=rel_conc), size=1, shape=15) +
  scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) + # A or F
  #scale_colour_viridis(option = 'H') +
  scale_y_reverse() +
  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
  #labs(colour = 'Log10 normalized abundance)', x = 'Month', y = 'Depth (m)') +
  labs(colour = '#/L', x = 'Month', y = 'Depth (m)')
  #theme(legend.title = element_text(angle = -90))
  #guides(fill = guide_legend(title.theme = element_text(size = 200)))
  #+ theme(legend.position = 'none')
  #geom_path(data = particles_bis, aes(juld, MLD), colour = "black") +
  #theme(legend.position = 'none') #+
  #geom_hline(yintercept = c(200,500,1000), linetype = 'dotted') #+
  #guides(colour = guide_legend(title.position = "bottom"))

#Normalized concentrations (#/L)

up_part

down_part <- particles_bis |> filter(depth < 2000, size >= 512) |>
  #group_by(size) |>
  ggplot() + facet_grid(wmo~size) + theme_bw() +
  geom_point(aes(x=juld, y=depth, colour=log10(rel_conc)), size=1, shape=15) +
  #geom_point(aes(x=juld, y=depth, colour=rel_conc), size=1, shape=15) +
  scale_colour_viridis(option = 'A', limits = c(-2.5,-1), oob = scales::squish, na.value = 'grey', direction = -1) +
  #scale_colour_viridis(option = 'H') +
  scale_y_reverse() +
  scale_x_date(labels = date_format("%m"), date_breaks = '3 month') +
  #labs(colour = 'Log10 normalized abundance)', x = 'Month', y = 'Depth (m)') +
  labs(colour = '#/L', x = 'Month', y = 'Depth (m)')

down_part

@@

### Processes at play

```{r, monthly anomaly between drifting at 500 and 1000 m}
#| label: fig-processes-at-play
#| fig-cap: Ratio of monthly mean particles concentrations during drifting (1000 m over 500 m) for size classes below 1.63 mm.

tmp <- uvp_data |> filter(park_depth != '200 m') |> group_by(wmo, size, month, park_depth) |> 
  summarize(monthly_mean_conc = mean(conc), monthly_median_conc = median(conc)) |>
  # compute concentration anomaly at 500 and 1000 m
  mutate(diff_anomaly = monthly_mean_conc - lead(monthly_mean_conc), div_anomaly = monthly_mean_conc/lead(monthly_mean_conc)) |> drop_na(diff_anomaly) |> #filter(is.finite(div_anomaly)) |>
  select(wmo, size, month, diff_anomaly, div_anomaly) |> mutate(month = month.abb[month]) |> 
  mutate(size = as.character(size)) |>
  mutate(rev_div_anomaly = 1/div_anomaly) |> 
  #filter(is.finite(rev_div_anomaly)) |>
  mutate(diff_anomaly = round(diff_anomaly, 1), rev_div_anomaly = round(rev_div_anomaly, 1), div_anomaly = round(div_anomaly, 1)) |>
  ungroup()

# test en weekly
# tmp <- uvp_data |> filter(park_depth != '200 m', month != 5) |> group_by(wmo, size, week, park_depth) |> 
#   summarize(monthly_mean_conc = mean(conc), monthly_median_conc = median(conc)) |>
#   # compute concentration anomaly at 500 and 1000 m
#   mutate(diff_anomaly = monthly_mean_conc - lead(monthly_mean_conc), div_anomaly = monthly_mean_conc/lead(monthly_mean_conc)) |> drop_na(diff_anomaly) |> #filter(is.finite(div_anomaly)) |>
#   select(wmo, size, week, diff_anomaly, div_anomaly) |> 
#   mutate(size = as.character(size)) |>
#   mutate(rev_div_anomaly = 1/div_anomaly) |> 
#   #filter(is.finite(rev_div_anomaly)) |>
#   mutate(diff_anomaly = round(diff_anomaly, 1), rev_div_anomaly = round(rev_div_anomaly, 1)) |>
#   ungroup()
# 
# create_data_for_heatmap <- function(float){
#   y <- unique(tmp$size)[1:12] # above, data is way too low
#   d <- tmp |> filter(wmo == float, size %in% y, week > 5)
#   x <- unique(d$week)
#   data <- expand.grid(X=x, Y=y)
#   #data$Z <- tt$diff_anomaly
#   data$Z <- d$rev_div_anomaly
#   data$wmo <- float
#   return(data)
# }

# d1 <- create_data_for_heatmap(6904241)
# d2 <- create_data_for_heatmap(1902578)
# d3 <- create_data_for_heatmap(4903634)
# d4 <- create_data_for_heatmap(6904240)
# 
# dd <- rbind(d1, d2, d3, d4)

# ggplot(dd, aes(X, Y, fill= Z)) + 
#   geom_tile() + scale_fill_viridis() + facet_wrap(~wmo)

create_data_for_heatmap <- function(float, df){
  y <- unique(df$size)[1:12] # above, data is way too low
  d <- df |> filter(wmo == float, size %in% y)
  #x <- c('Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan')
  x <- unique(d$month)
  data <- expand.grid(X=x, Y=y)
  #data$Z <- tt$diff_anomaly
  data$Z <- d$rev_div_anomaly
  #data$Z <- d$div_anomaly
  data$wmo <- float
  return(data)
}

# for some reason, there is some bug with the quarto compilation .. so I save the tibble .. and reload it 
# d1 <- create_data_for_heatmap(6904241, tmp)
# d2 <- create_data_for_heatmap(1902578, tmp)
# d3 <- create_data_for_heatmap(4903634, tmp)
# d4 <- create_data_for_heatmap(6904240, tmp)
# dd <- rbind(d1, d2, d3, d4)
# vroom_write(dd, 'heatmap_data.csv')

# reload + redo some annoying stuff that were lost during the writing of the csv
dd <- vroom('heatmap_data.csv') |> mutate(Y = as.character(Y)) 
dd$X <- factor(dd$X, levels = c('May', 'Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan'))
dd$Y <- factor(dd$Y, levels = c('102', '128', '161', '203', '256', '323', '406', '512', '645', '813', '1020', '1290'))
dd$Y <- factor(dd$Y, levels = c('1290', '1020', '813', '645', '512', '406', '323', '256', '203', '161', '128', '102'))
dd$wmo <- factor(dd$wmo, levels = c(6904240,4903634,6904241,1902578))

#dd <- dd |> mutate(Z_binned = cut(Z, breaks = c(0,1,1.5,2.5,max(Z))))
dd <- dd |> mutate(Z_binned = cut(Z, breaks = c(0,.5,1,1.5,2,max(Z))))

# ggplot(dd) +
#   geom_tile(aes(X, Y, fill = Z), alpha = 0.75) + scale_fill_viridis() + theme_bw() + labs(x = 'Month', y = 'Size class') +
#   facet_wrap(~wmo) +
#   geom_text(data = dd, aes(X, Y, label = Z))

ggplot(dd) + 
  geom_tile(aes(X, Y, fill = Z_binned), colour = 'black', alpha = 1) + theme_bw() + labs(fill = 'Ratio', x = 'Month', y = 'Size class') +
  facet_wrap(~wmo) + scale_fill_viridis_d(option = 'A')

# plot_heatmap <- function(float){
#   y <- unique(tmp$size)[1:12]
#   tt <- tmp |> filter(wmo == float, size %in% y) 
#   x <- c('Jun', 'Jul', 'Aug', 'Sep', 'Oct', 'Nov', 'Dec', 'Jan')
#   data <- expand.grid(X=x, Y=y)
#   #data$Z <- tt$anom
#   data$Z <- tt$diff_anomaly
#   data$Z <- tt$rev_div_anomaly
# 
#   p <- ggplot(data, aes(X, Y, fill= Z)) + 
#     geom_tile() + scale_fill_viridis() +
#     geom_text(data = data, aes(X, Y, label = Z)) +
#     labs(title = float)
#   
#   return(p)
# }
# 
# p1 <- plot_heatmap(4903634)
# p2 <- plot_heatmap(6904240)
# p3 <- plot_heatmap(1902578)
# p4 <- plot_heatmap(6904241)
# 
# (p1 | p2) / 
# (p3 | p4)

@@

```{r, test gt table}

monthly_mean <- uvp_data |> mutate(month = month.abb[month(juld)]) |> group_by(size, park_depth, month) |> summarize(mean_conc = round(mean(conc, na.rm=T), 2)) |> ungroup()

# pivot wider
monthly_mean <- monthly_mean |> pivot_wider(names_from = month, values_from = mean_conc) |> select(size, depth = park_depth, May, Jun, Jul, Aug, Sep, Oct, Nov, Dec, Jan)

monthly_mean |> group_by(size) |> gt(rowname_col = 'depth')

#monthly_mean |> gt(rowname_col = c('size')) |> tab_stubhead(label = 'Class')
